Parsing Postfix log files

### 2007-05-14 doc/logparser.tex
Documented and discussed identifying bounce notifications.  It'll need a second draft, but this is the basis of it.

>>>>

\subsection{XXX}

\subsubsection{Identifying bounce notifications}

\label{identifying-bounce-notifications}

Postfix 2.2.x (and presumably previous versions) lacks explicit logging
when bounce notifications are generated; suddenly there will be log entries
for a mail which lacks an obvious source.  There are similarities to the
problem of child mails discussed in section
\ref{tracking-re-injected-mail}, but unlike the solution described therein
bounce notifications do not eventually have a log line which identifies
their source.  The solution is that heuristics must be used to identify
them, and those heuristics are:

    \item The sender address will be <>.

    \item Neither \texttt{postfix/smtpd} nor \texttt{postfix/pickup} will
        have logged any messages associated with the mail, viz it was
        generated internally by Postfix, not accepted via SMTP or submitted
        locally by sendmail.

    \item The mail will not have been tagged as a tracked mail (see section
        \ref{tracking-re-injected-mail}).

    \item The message-id has a specific format: \newline
        \texttt{YYYYMMDDhhmmss.queueid@server.hostname} \newline
        e.g. \texttt{20070321125732.D168138A1@smtp.cs.tcd.ie}

        The message-id is generated by Postfix for this bounce
        notification, so the queueid in the message-id must be the same as
        the queueid of the mail: this is what distinguishes bounce
        notifications generated locally from bounce notifications which are
        being forwarded inwards as a result of aliasing.  In the later case
        the message-id will be unchanged, and so even if it happens to be
        in the correct format (e.g. if it was generated by Postfix on
        another server) the queueid in the message-id will not match the
        queueid of the mail.

Once a mail has been identified as a bounce notification the unknown origin
flag is cleared and the mail can be entered in the database.

There is a small chance that a mail will be incorrectly identified as a
bounce notification, as the heuristics used may be too broad.  For this to
occur the following conditions would have to be met:

    \item The mail must have been generated internally by Postfix.

    \item The sender address must be <>.

    \item The message-id must have the correct format and match the queueid
        of the mail.  Whist it is possible that a mail sent from elsewhere
        could easily have the correct message-id format, the chance that
        the queueid in the message-id would match the queueid of the mail
        is extremely small.

The most likely cause of mis-identification is if a mail generated
internally by Postfix is identified as a bounce notification when it is a
different type of message; arguably this is a benefit rather than a
drawback, as future mails generated internally by Postfix will be handled
correctly.  XXX FIND OUT IF DSNS FIT THE PATTERN.

\subsubsection{Aborted delivery attempts}

\label{aborted-delivery-attempts}

\subsubsection{Timeouts during DATA phase}

\label{timeouts-during-data-phase}

\subsubsection{Out of order log entries}

\label{out-of-order-log-entries}
